Historical Soil Organic Carbon Budget

### 2020-03-26 SOCbudgetPaper/SOCbudget.tex
update residue part (and before)

\hypertarget{sec:agrimanagement}{%
\subsection{Agricultural management (50)}\label{sec:agrimanagement}}

@@

Our land use patterns are based on the Land-Use Harmonization 2 (LUH2,
\citep{LUH2}) data set, which we aggregate from quarter degree to half
degree resolution. To transform the five different cropland
subcategories of LUH2 into our crop functional types we define country
and year specific area shares based on FAOSTAT data (\citep{FAOSTAT})
(see @ref(appendix:fao2luh2mag). Where no data is available we fill data
gaps by using similar countries as proxies or global averages. Land use
transitions are calculated as area differences of the land use data,
allowing any land use type to either expand or shrink in each time step.

\subsubsection{Crop Production and Residues (300)}

Using half-degree yield data from LPJmL (\citep{LPJmL4_1}) as well as
half-degree cropland pattern (see @ref(\#sec:landuse)) we compile crop
functional type specific half-degree production patterns. We calibrate
them to country level FAOSTAT production (\citep{FAOSTAT}) values to be
in line with historical observations.

Starting from this harvested fraction we calculate above ground and
below ground residual biomass using harvest indices and shoot:root
ratios as described in (\citep{bodirsky2012}). We assume that all below
ground biomass is left on the fields, whereas above ground biosmass can
be drawn away for other purposes such as feeding animals
(\citep{weindl}), fuel or for material uses. We calculate recycle shares
by subtracting all residue usages (see \citep{bodirsky1012}) from the
total above ground biomass.

To transform dry matter values into carbon quantities used as soil
input, we compiled crop functional type specific carbon to dry matter
(c:dm) ratios for harvested organs and applied a universal c:dm ratio of
0.44 for all above and below ground residue biomass.

\subsubsection{Pasture production}

Using livestock production statistics as well as feed mix assumptions as
describted in (\citep{weindl}) we estimating country specific pasture
production.

Using the same method as for crop production we disaggregate and
calibrate half-degree pasture production pattern.

@@

Since their is no native, biophysical indicator to disaggregate country
level FAOSTAT livestock production values to half-degree pattern, we use
the following rule based assumptions:

\begin{itemize}
\item
  Based on the development state of a country, we devide poultry, egg
  and monogastric meat production into intensive and extensive systems.
\item
  Extensive production is allocated to urban area share within a country
  based on the assumption that these animals are held in households,
  subsistence or small-holder farming systems that are labour rich.
\item
  Intensive production is distributed within a country using the crop
  production (excluding 2nd generation bioenergy crops) share, assuming
  that feed availability is the most driving factor for livestock
  location.
\item
  We split dairy and ruminated meat production into pasture fed and
  non-pasture fed fractions based on the aggregated feed mix within a
  country
\end{itemize}

\begin{itemize}
\item
  Pasture fed animals are apportioned to half-degree cells according to
  the pasture production shares of that cell within a country.
\item
  Non-pasture fed animals are distributed again using crop production
  (excluding 2nd generation bioenergy crops) shares.
\end{itemize}

We calculate excretions using the nitrogen balance of the livestock
system (\citep{bodirsky1012}) by assuming that every feed intake not
incoorpared in the slaugher mass, has to excreted. Using C:N ratios
(given by \citep[(][]{ipcc_2019_2019}) we estimate carbon content

@@

Simple growing period calculations together with irrigation shares of
LUH2v2 are used (BB: you dont like d's, on't you?) to estimate water
effects on decay rates.

>>>>

\hypertarget{sec:agrimanagement}{%
\subsection{Agricultural management data on 0.5 degrees
(50)}\label{sec:agrimanagement}}

@@

Land use patterns are based on the Land-Use Harmonization 2 (LUH2,
\citep{LUH2}) data set, which we aggregate from quarter degree to half
degree resolution. We disaggregate the five different cropland
subcategories of LUH2 into our 17 crop groups, assuming relative shares
for each gridcell based on the country and year specific area shares of
FAOSTAT data (\citep{FAOSTAT}). Land use transitions are calculated as
net area differences of the land use data on half-degree.

\subsubsection{Crop, Crop Residues and Pasture Production (300)}

Using half-degree yield data from LPJmL (\citep{LPJmL4_1}) as well as
half-degree cropland patterns (see @ref(\#sec:landuse)) we compile crop
functional type specific half-degree production patterns. We calibrate
cellular yields with one country-level calibration factor for each
crop-group to meet historical FAOSTAT production (\citep{FAOSTAT}).

Crop residues production is based on a revised methodology as described
in (\citep{bodirsky2012}) and will be explained in key aspects again due
to its central role for soil carbon modelling. Starting from crop
production estimates of the harvested organs and their respective crop
area, we estimate above-ground (ag) and below-ground (bg) residual
biomass using yield-dependent harvest indices and shoot:root ratios. We
assume that all bg residues are recycled to the soil, whereas ag
residues can be burned or harvested for other purposes such as feeding
animals (\citep{weindl}), fuel or for material use.

A fixed share of the ag residues is assumed to be burned on field, which
depends on the per-capita income of the country. Following
\citep{smil1999}) we assume 25\% burn share for low-income countries
according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr}\)), 15\%
for high-income (\(>\,10000\,\tfrac{USD}{yr}\) and linearly interpolated
shares for all middle-income countries.

Residue demand for feed is based on regional livestock specific feed
basekts (see \citep{weindl}) taking available ag residual biomass as
well as livestock production systems and thus again development state
into account. Here we aggregate crop groups to residue groups based on
their suitability as feed (see @ref(append:Table\_kcr2kres)).

We estimated a material use share for the straw residues group of 5\%
and a fuel demand share demand of 10\% in low income countries.

The remaining ag residues are assumend to be recycled to the soil.

Using livestock production statistics as well as feed mix assumptions as
describted in (\citep{weindl}) we estimating country specific pasture
production. Following the same approach as for crop production we
disaggregate and calibrate half-degree pasture production pattern from
grass yields from LPJmL and pasture area and rangeland patterns ( (see
@ref(\#sec:landuse))) to derive half-degree pasture production patterns.

To transform dry matter estimates into carbon, we compiled crop and
residue group specific carbon to dry matter (c:dm) ratios (see appendix
table X) {[}citations{]} for harvested organs and applied a universal
c:dm ratio of 0.44 {[}LPJml???{]} for all above and below ground residue
biomass as well as for grass biomass.

@@

To disaggregate country level FAOSTAT livestock production values to
half-degree pattern, we use the following rule based assumptions which
was inspired by the approach of \citep{gilbert}.

For poultry, egg and monogastric meat production we

\begin{itemize}
\item
  divide into intensive and extensiv eproduction systems based on a
  per-capita (explain rules in more detail)
\item
  Based on the development state (thats not a fixed term) of a country,
  we devide poultry, egg and monogastric meat production into intensive
  and extensive systems.
\item
  Extensive production is allocated to built-up area share within a
  country based on the assumption that these animals are held in
  households, subsistence or small-holder farming systems.
\item
  Intensive production is distributed within a country using the crop
  production (excluding 2nd generation bioenergy crops) share, assuming
  that feed availability is the most driving factor for livestock
  location.
\end{itemize}

For dairy and ruminated meat production we split into pasture fed and
non-pasture fed fractions based on the aggregated feed mix within a
country (explain rules)

\begin{itemize}
\item
  Pasture fed animals are distributed to half-degree cells according to
  the pasture production shares of that cell within a country (not
  methodologically clear description).
\item
  Non-pasture fed animals are distributed again using crop production
  (excluding 2nd generation bioenergy crops) shares (not clear, either).
\end{itemize}

We calculate excretions by first estimating the nitrogen balance of the
livestock system (\citep{bodirsky1012}) on the basis of comprehensive
livestock feed baskets ({[}weindl{]}), assuming that all nitrogen in
protein feed intake, minus the nitrogen in the slaugher mass, is
excreted. Carbon in excreted manure is estimated by applying fixed C:N
ratios (given by \citep[(][]{ipcc_2019_2019}). What about animal waste
management losses?

@@

Simple growing period calculations together with irrigation shares of
LUH2v2 are used (BB: you dont like d's, on't you) to estimate water
effects on decay rates.

### 2020-03-26 SOCbudgetPaper/methods.Rmd
update residue part (and before)

Land use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution. We disaggregate the five different cropland subcategories of LUH2 into our X crop groups, assuming relative shares for each gridcell based on the country and year specific area shares of FAOSTAT data ([@FAOSTAT]) (see \@ref(appendix:fao2luh2mag)<!--glaube nicht dass du da noch n appendix brauchst-->. For X countries without FAOSTAT data, we fill data gaps by using similar countries as proxies or global averages<!--heisst das die haben HURTT data aber keine FAO daten? woher kriegt dann hurtt die daten?-->. Land use transitions are calculated as area differences of the land use data.

### Crop Production and Residues	(300)
Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(#sec:landuse)) we compile crop functional type specific half-degree production patterns. We calibrate cellular yields with one country-level calibration factor for each crop-group to meet historical FAOSTAT production ([@FAOSTAT]).

Starting from these production estimates of the harvested organs and their respective crop area, we estimate above-ground and below-ground residual biomass using yield-dependent harvest indices and shoot:root ratios as described in ([@bodirsky2012]). We assume that all below ground biomass is left on the fields. A fixed share of the above-ground residues is assumed to be burnt on field, which depends on the per-capita income of the country. For low-income countries according to worldbank definitions (<1000 USD/yr), we assume a ... etc.
ground biomass can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material uses. We calculate recycle shares by subtracting all residue usages (see [@bodirsky1012, @weindl2019]) from the total above ground biomass.

<!-- Add formular or drawings here? --> 
<!-- if you want, but first finish the rest -->
<!-- Ask benni if it is ok to just cite his paper here and not go into detail, how we distribute the residues --> <!-- Benni: No, explain in detail as it is important for your manuscript and adds credibility to your detailed calcualtions. Also, some assumptions may have changed (e.g. DevelopmentState as function of gdp) -->

To transform dry matter estimates into carbon, we compiled crop functional type specific carbon to dry matter (c:dm) ratios (see appendix table X) [citations] for harvested organs and applied a universal c:dm ratio of 0.44 [citation] for all above and below ground residue biomass.

<!-- Do we have to explain our choice for the crop functional type specific data. Anyway we have to discuss here a bit, because the overall balance is not clear to me. -->

### Pasture production
Using livestock production statistics as well as feed mix assumptions as describted in ([@weindl]) we estimating country specific pasture production. 
<!-- Is this true? Ask Isabelle again. -->

Using the same method as for crop production we disaggregate and calibrate half-degree pasture production pattern.
<!--As for cropland production we used pasture yields from LPJmL and pasture and rangeland patterns to derive half-degree pasture production patterns. Again country specific pasture production is used to calibrated the pattern.-->

@@

For poultry, egg and monogastric meat production we 
* divide into intensive and extensiv eproduction systems based on a per-capita (explain rules in more detail)

@@

For dairy and ruminated meat production we

* splot into pasture fed and non-pasture fed fractions based on the aggregated feed mix within a country (explain rules)

>>>>

Land use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution. We disaggregate the five different cropland subcategories of LUH2 into our 17 crop groups, assuming relative shares for each gridcell based on the country and year specific area shares of FAOSTAT data ([@FAOSTAT]). Land use transitions are calculated as net area differences of the land use data on half-degree.

### Crop, Crop Residues and Pasture Production	(300)

Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland patterns (see \@ref(#sec:landuse)) we compile crop functional type specific half-degree production patterns. We calibrate cellular yields with one country-level calibration factor for each crop-group to meet historical FAOSTAT production ([@FAOSTAT]).

Crop residues production is based on a revised methodology as described in ([@bodirsky2012]) and will be explained in key aspects again due to its central role for soil carbon modelling. Starting from crop production estimates of the harvested organs and their respective crop area, we estimate above-ground (ag) and below-ground (bg) residual biomass using yield-dependent harvest indices and shoot:root ratios. We assume that all bg residues are recycled to the soil, whereas ag residues can be burned or harvested for other purposes such as feeding animals ([@weindl]), fuel or for material use.

A fixed share of the ag residues is assumed to be burned on field, which depends on the per-capita income of the country. Following [@smil1999]) we assume 25% burn share for low-income countries according to worldbank definitions ($<\,1000\,\tfrac{USD}{yr}$), 15% for high-income  ($>\,10000\,\tfrac{USD}{yr}$ and linearly interpolated shares for all middle-income countries.

Residue demand for feed is based on regional livestock specific feed basekts (see [@weindl]) taking available ag residual biomass as well as livestock production systems and thus again development state into account. Here we aggregate crop groups to residue groups based on their suitability as feed (see \@ref(append:Table_kcr2kres)).

We estimated a material use share for the straw residues group of 5% and a fuel demand share demand of 10% in low income countries.

The remaining ag residues are assumend to be recycled to the soil.

<!-- Add formular or drawings here? --> 
<!-- if you want, but first finish the rest -->

Using livestock production statistics as well as feed mix assumptions as describted in ([@weindl]) we estimating country specific pasture production. Following the same approach as for crop production we disaggregate and calibrate half-degree pasture production pattern from grass yields from LPJmL and pasture area and rangeland patterns ( (see \@ref(#sec:landuse))) to derive half-degree pasture production patterns.

To transform dry matter estimates into carbon, we compiled crop and residue group specific carbon to dry matter (c:dm) ratios (see appendix table X) [citations] for harvested organs and applied a universal c:dm ratio of 0.44 [LPJml???] for all above and below ground residue biomass as well as for grass biomass.

@@

For poultry, egg and monogastric meat production we

* divide into intensive and extensiv eproduction systems based on a per-capita (explain rules in more detail)

@@

For dairy and ruminated meat production we split into pasture fed and non-pasture fed fractions based on the aggregated feed mix within a country (explain rules)
